---
title: "Data analysis skeleton: Update to reflect your team's analyses!"
output: 
  html_document:
    keep_md: true
---

# Load necessary packages
```{r}
#install.packages('pacman', repos = "http://cran.us.r-project.org")
pacman::p_load(tidyverse, broom, lubridate, gvlma, car, ggpubr, MASS)
```

# Load cleaned data files
## Data type 1: Small mammals
```{r}
mammals <- read_csv('./clean_data/mammals_clean.csv') %>% 
  mutate(year = year(collectDate),
         month = month(collectDate),
         date = make_date(year, month),
         tagID = factor(tagID)) %>%
  dplyr::select(siteID, year, month, date, plotID, taxonID, tagID, adultTicksAttached)
```

### Summarize mammals to monthly counts by focal taxa and overall total
```{r}
mammals_monthly <- mammals %>%
  group_by(siteID, year, month, taxonID) %>%
  summarize(mammal_count = n()) %>% 
  mutate(date = make_date(year, month)) %>%
  dplyr::select(siteID, year, month, date, taxonID, mammal_count) %>% 
  pivot_wider(names_from = taxonID, values_from = mammal_count) %>%
  ungroup() %>% 
  mutate(total_mammals = rowSums(across(PEGO:RARA), na.rm=T)) %>%
  dplyr::select(siteID:date, total_mammals, PEGO:RARA)
  #filter(taxonID=="PELE")
```

## Data type 2: Air temp
```{r}
temp <- read_csv('./clean_data/temp_clean.csv')
```

### Summarize temp to monthly mean
```{r}
temp_monthly <- temp %>%
  group_by(siteID, year, month) %>%
  summarize(monthlyTemp = round(mean(tempSingleMean),2)) %>%
  mutate(date = make_date(year, month)) %>%
  dplyr::select(siteID, year, month, date, monthlyTemp)
```

## Data type 3 Ticks
```{r}
ticks <- read_csv('./clean_data/ticks_clean.csv') %>% 
  mutate(year = year(collectDate),
         month = month(collectDate),
         date = make_date(year, month)) %>%
  dplyr::select(siteID, year, month, date, adultCount:larvaCount)
```

### Summarize ticks to monthly total
```{r}
ticks_monthly <- ticks %>% 
  group_by(siteID, date) %>% 
  summarize(sum_adult=sum(adultCount),
            sum_nymph= sum(nymphCount),
            sum_larva=sum(larvaCount))
```


## Combine data files based on shared criteria
Type out some details of what the column names are that your data objects have in common. Note that for R to combine them, the column names must be **exactly** the same (e.g., spelling, capitalization, etc.)
```{r}
monthly <- full_join(temp_monthly, ticks_monthly) %>%
  full_join(., mammals_monthly)

print(monthly)
```

# Statistical analysis:
## Summary/descriptive statistics
Often, we report summary statistics like the mean, a measure of variance (standard deviation, standard error, 95% confidence intervals) and the sample size. 

We can do this easily through the use of `dplyr()` commands. In this example, we can calculate average annual values per site using:
```{r}
temp_summary_by_site <- monthly %>% 
  group_by(siteID) %>% 
  summarize(n = n(),  # Sample size
            mean_Temp = mean(monthlyTemp, na.rm=T), sd_Temp = sd(monthlyTemp, na.rm=T), se_Temp = sd_Temp/sqrt(n)) # Mean, sd, se for temperature

glimpse(temp_summary_by_site)
```

To simplify reporting, we can also ask R to round all of these calculated values to a specific number of decimal places:
```{r}
temp_summary_by_site <- as_tibble(temp_summary_by_site) %>% 
  mutate_at(vars(mean_Temp:se_Temp), funs(round(.,2))) %>% # Round each of our calculated columns to 2 decimals
  arrange(mean_Temp) 

print(temp_summary_by_site)
```

# ANOVA: Do monthly adult tick counts vary among sites?
First, we'll define our ANOVA model:
```{r}
adults_per_site <- aov(sum_adult ~ siteID,  data= monthly )
```

Next, we can use this quick command to check the ANOVA model assumptions:
```{r, warning=F}
# Check for homogeneity of variance: 
leveneTest(sum_adult ~ siteID,  data= monthly)
# If the p-value of Levene's test is greater than 0.05, we have not violated the assumption of homogeneity of variance

# Check for normality of residuals:
## 1) Extract the model residuals
aov_residuals <- residuals(object = adults_per_site)

## 2) Run Shapiro-Wilk test
shapiro.test(x = aov_residuals )

```

*Since our data violate the homogeneity of variance assumption, we instead do a Welch one-way test using the `oneway.test()` function*
```{r}
oneway.test(sum_adult ~ siteID,  data= monthly)

pairwise.t.test(monthly$sum_adult,
                monthly$siteID,
                 p.adjust.method = "BH", pool.sd = FALSE)
```
If the model p-value is smaller than 0.05 (our standard alpha), we reject the null hypothesis and conclude that there is a significant difference in our dependent variable between *at least two* sites. 

We can then look at the pairwise comparison table to see which sites differed from one another and update our figure to reflect what we found

# Box plot: Tick counts among Sites
```{r ticks_per_site}
ggplot(data = monthly,aes(x = siteID,
                           y = sum_adult)) +
  geom_jitter(width = 0.05, alpha = 0.5) + # Add partially-transparent data points 
  geom_boxplot(aes(fill = siteID), outlier.color = NA, 
               alpha = 0.5) + # Plot boxplots
  theme_minimal() +
  scale_fill_manual(values = c("darkorange","purple","cyan4", "firebrick2", "springgreen1"), # Set colors for groups
                    guide = 'none') + # Turn off legend for this layer
  labs(x = "NEON site", # Specify x-axis label
       y = "Adult ticks per month") + # Specify y-axis label
  annotate("text", x = 1:5, y = c(6, 32,6,62,32), label = c("A","B", "A","C","B")) # Add a text label to differentiate the statistically-different groups
``` 

# Correlation & Linear regression: Monthly tick counts as a function of mean temperature
**We are comparing our monthly temperature data to tick populations across different sites.**

We can calculate a correlation coefficient to quantify the strength and direction of association between the two variables.

The types of correlation you can use are:  

* Pearson: Parametric correlation (requires data be normally distributed) for continuous data   

* Spearman: Non-parametric (e.g., if your data are not normally distributed) correlation can also take ordinal data

```{r}
# Check whether tick data normally distributed
shapiro.test(monthly$sum_adult)

# Check whether monthly temperature data normally distributed
shapiro.test(monthly$monthlyTemp)

```

We will use Spearman correlation because the data are not normally distributed
```{r}
cor_alpha <- cor.test(monthly$sum_adult, 
                      monthly$monthlyTemp, 
                      method = "spearman")

tidy(cor_alpha)
```

Now, let's set up and run a model to conduct a regression.
First, we can run a linear regression model and check its assumptions:
```{r, warning=F}
tempVticks <- lm(sum_adult ~ monthlyTemp, data = monthly)
gvlma(tempVticks)
```
Our linear model **failed** the overall model assumptions, so we should not go forward interpreting it.

We will instead use a **negative binomial regression** because our response variable includes count data, and is overdispersed (the variances is much larger than the mean) 
```{r}
tempVticks <- glm.nb(sum_adult ~ monthlyTemp, data = monthly)
summary(tempVticks)
car::Anova(tempVticks)

# Report the chi-square as X2 (degress of freedom, N = sample size) = chi-square statistic value, p = p value.
```

# Scatterplot: Monthly adult ticks as a function of monthly mean temperature
Remember that ecological data reporting convention is to **omit** the trend line for non-significant relationships.

```{r ticks_vs_temp_scatterplot}
ggplot(data = monthly,aes(x = monthlyTemp,
                           y = sum_adult)) +
  geom_point(aes(color = siteID), # Add data points
             size = 2) +
  theme_minimal() +
  scale_x_continuous(limits = c(0, 30)) + # Set min/max x-axis values
  scale_y_continuous(limits = c(0, 60), # Set min/max y-axis values
                     breaks = seq(0, 60, 10)) + # Specify axis break values
  scale_color_manual(values = c("darkorange","purple","cyan4", "firebrick2", "springgreen1")) + # Specify custom colors
  labs(x = "Monthly Average Temperature (°C)", # Specify axis labels
       y = "Number of Adult Ticks", 
       color = "NEON site")
```

# Linear regression (Mammals vs Ticks)
**We are comparing the small mammal populations with tick populations at the sites.**
```{r}
# Check whether tick data normally distributed
shapiro.test(monthly$sum_adult)

# Check whether monthly temperature data normally distributed
shapiro.test(monthly$total_mammals)

```

We will use Spearman correlation because the data are not normally distributed
```{r}
cor_alpha <- cor.test(monthly$sum_adult, 
                      monthly$total_mammals, 
                      method = "spearman")

tidy(cor_alpha)
```

Now, let's set up and run a model to conduct a regression.

First, we can run a linear regression model and check its assumptions:
```{r, warning=F}
mammVticks <- lm(sum_adult ~ total_mammals, data = monthly)
gvlma(mammVticks)
```
Our linear model **failed** the overall model assumptions, so we should not go forward interpreting it.

We will instead use a **negative binomial regression** because our response variable includes count data, and is overdispersed (the variances is much larger than the mean) 
```{r}
mammVticks <- glm.nb(sum_adult ~ total_mammals, data = monthly)
summary(mammVticks)
car::Anova(mammVticks)
```

# Scatterplot: Monthly adult ticks as a function of monthly mammals
```{r ticks_vs_mammals_scatterplot}
ggplot(data = monthly, aes(x = total_mammals, # name the variable that goes on the x-axis
                           y = sum_adult,  # name the variable that goes on the y-axis
                           col = siteID)) + # name the variable to differentiate colors
  geom_point(pch = 16, size = 3) +  # Graph these variables as a scatterplot
  theme_minimal() +
  scale_x_continuous(limits = c(0, 350)) + # Set min/max x-axis values
  scale_y_continuous(limits = c(0, 60), # Set min/max y-axis values
                     breaks = seq(0, 60, 10)) + # Specify axis break values
  scale_color_manual(values = c("darkorange","purple","cyan4", "firebrick2", "springgreen1")) + # Specify custom colors
  labs(x = "Small mammals captured", # Specify axis labels
       y = "Adult Ticks", 
       color = "NEON Site")+
  geom_smooth(aes(group=1), method = "glm.nb", col='black')
```
